---
title: "Fall 2019 - Data 607 - Week 3 Assignment"
author: "Charlie Rosemond - 091519"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>

I started by loading stringr to facilitate analysis.

```{r load}
library(stringr)
```

<br>

### 3. Copy the introductory example.

I extracted the names, identified by at least 2 sequences of alphabetic characters, from the character string.

```{r names, eval=TRUE}
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
names <- unlist(str_extract_all(raw.data, "[[A-z].,' ']{2,}"))
names
```

<br>

#### 1. Use the tools of this chapter to rearrange the vector so that all elements conform to the standard "first_name last_name".

I copied the list over then identified and replaced the relevant strings, using my knowledge of The Simpsons to change "C." to "Charles". I then used the word function in stringr to extract the first and last names, which I displayed as columns in a data frame.
    
```{r number31, eval=TRUE}
names1 <- names
grep(",", names1, value = TRUE)
names1 <- str_replace(names1, "Burns, C. Montgomery", "Charles Montgomery Burns")
names1 <- str_replace(names1, "Simpson, Homer", "Homer Simpson")
grep(".", names1, value = TRUE)
names1 <- str_replace(names1, "[A-z]{2,3}\\.+\\s", "")
names1
first_name <- word(names1, 1)
last_name <- word(names1, -1)
data.frame(first_name, last_name)

```

<br>

#### 2. Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).
    
I located persons with titles by using a pattern of 2 or 3 alphabetic characters preceding a period. This pattern avoided the "C." initial for Mr. Burns.

```{r number32, eval=TRUE}
titles <- grepl("[A-z]{2,3}\\.", names)
titles
data.frame(names, titles)
```

<br>

#### 3. Construct a logical vector indicating whether a character has a second name.
    
Using the cleaned list names1, I created an object containing the string counts for each person. I then created a logical vector indicating TRUE for each person with more than 2 strings, or names.
    
```{r number33, eval=TRUE}
second_name_check <- str_count(names1, "[\\w+]{3,}")
second_name <- second_name_check > 2
second_name
data.frame(names1, second_name_check)
```

<br>

### 4. Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

<br>

#### 1. "[0-9]+\\$"
    
This regular expression would match a string consisting of some number of numeric digits ending in a dollar sign. 
    
```{r number41, eval=TRUE}
one <- c("abcdef$", "123456", "abcdef", "123456$")
grepl("[0-9]+\\$", one)
```

<br>

#### 2. "\\b[a-z]{1,4}\\b"
    
This regular expression would match a string consisting of one to four lowercase letters with an empty string on both edges.
    
```{r number42, eval=TRUE}
two <- c("Charlie", "Name", "charlie", "one", "four")
grepl("\\b[a-z]{1,4}\\b", two)
```

<br>

#### 3. ".*?\\.txt$"
    
This regular expression would match a string ending in ".txt" (e.g. a .txt file).
    
```{r number43, eval=TRUE}
three <- c("some.txt", "file.csv", "another.txt", "other.xls")
grepl(".*?\\.txt$", three)
```

<br>

#### 4. "\\d{2}/\\d{2}/\\d{4}"
    
This regular expression would match a string consisting of, in order, two digits, a forward slash, two digits, a forward slash, and four digits (e.g. a date).
    
```{r number44, eval=TRUE}
four <- c("9/12/19", "09/12/19", "09/12/2019")
grepl("\\d{2}/\\d{2}/\\d{4}", four)
```

<br>

#### 5. "<(.+?)>.+?</\\1>"

This regular expression appears to match pairs of properly formatted HTML tags and whatever they contain.

```{r number45, eval=TRUE}
five <- c("<html><body>Some text</body></html>", "</html>something</body>")
grepl("<(.+?)>.+?</\\1>", five)
```

<br>

### 9. The following code hides a secret message. Crack it with R and regular expressions.

I had difficulty finding the snippet from the book online, so I actually typed it out. In doing so, I figured the secret message related to either the numbers or the upper case letters (I wasn't sure what to make of "clcop"). Extracting uppercase letters spelled out a phrase, and upon closer inspection, the words in the phrase were separated by punctuation. I replaced the periods with spaces, creating "CONGRATULATIONS YOU ARE A SUPERNERD!"

```{r number9, eval=TRUE}
snippet <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
str_extract_all(snippet, "[0-9]")
str_extract_all(snippet, "[A-Z]")
uppercase <- unlist(str_extract_all(snippet, "[[A-Z].|!]"))
secret_message <- str_replace_all(paste(uppercase, collapse = ""), "\\.", " ")
secret_message
```

<br>

##### Sources:
"Basic Regular Expressions in R, Cheat Sheet". RStudio. Accessed 091219 from https://www.rstudio.com/wp-content/uploads/2016/09/RegExCheatsheet.pdf;

"Extract first word from a column and insert into new column [duplicate]". StackOverflow. Accessed 091219 from https://stackoverflow.com/questions/31925811/extract-first-word-from-a-column-and-insert-into-new-column;

"Chapter 8". Automated Data Collection with R. Accessed 091219 from http://kek.ksu.ru/eos/WM/AutDataCollectR.pdf
